---
title: "Bedgraph plots on genome for Ssd1 CRAC "
author: Edward Wallace
date: 2 October 2020
output:
  html_document:
    toc: true
    toc_depth: 4
---

# Summary

This script produces genome browser-style figures of SSD1 CRAC profiles from bedgraph data. It focuses on 30°C rep A (`Ssd1_3_30`) and B (`Ssd1_4_30`), showing profiles on extended full-length transcripts.

This script relies on bedgraph files for both plus and minus strands, produced by `genomeCoverageBed` while running the pipeline, put in the directory `/Ssd1_CRAC_demult_dedup_20190114_all/bedgraph_genomecov/`. So it must be run after  `src/CRAC_pipeline_SE_demult_dedup.py`, see the repository `README.md` for details.


### Load Packages

```{r setup,message=FALSE,echo=FALSE,cache=FALSE}
knitr::opts_chunk$set(warning=FALSE,message=FALSE,echo=FALSE,cache=TRUE,
                      results="hide",
                      fig.path="figure/bedgraph-",
                      cache.path="cache/bedgraph-")
# load common functions
library(tidyverse)
library(cowplot)
library(valr) # for read_bedgraph
theme_set(theme_cowplot(font_size = 11) + 
              theme(strip.background = element_blank(),
                    strip.text.y = element_text(angle=0)))

# setwd("~/Repos/Ssd1_CRACanalysis_2020/rmarkdown/")
```

```{r plot_functions}

onebreak <- function(lims,digits=1) {
    # select axis breaks/ticks with 0 and one tick
    c(0, signif(max(lims*.8), digits=digits) ) 
}

plot_bg <- function(bgbig,chromr,startr,endr,strandrev=FALSE,title=NULL,colourmain="blue") {
    # plot bedgraph in genomic co-ordinates
    bgplot <- 
        ggplot(data=bgbig %>%
                   filter(chrom==chromr, start >= startr, end <= endr),
               aes(ymin=0,ymax=value,xmin=start,xmax=end)) +
        geom_rect(colour=colourmain,fill=colourmain) +
        facet_grid(Type+Sample ~.,scales="free_y") +
        scale_y_continuous(expand=c(0,0),breaks=onebreak) +
        expand_limits(x=c(startr,endr)) + 
        labs(title=title, x=chromr, y="Reads")
    if (strandrev) {
        return( bgplot + scale_x_reverse(expand=c(0,0),limits=c(endr,startr)) )
    } else {
        return( bgplot + scale_x_continuous(expand=c(0,0),limits=c(startr,endr)) )
    }
}

read_gff <- function(file){
    # tidyverse read gff function from rmonad vignette
    # https://cran.r-project.org/web/packages/rmonad/vignettes/gff-processing.html
    readr::read_tsv(
        file,
        col_names = c(
            "chrom",
            "source",
            "type",
            "start",
            "end",
            "score",
            "strand",
            "phase",
            "attr"
        ),
        na        = ".",
        comment   = "#",
        col_types = "ccciidcic"
    )
}

plot_mRNA <- function(gff,gene,pad=0,gff_extra=NULL,colour_extra="red") {
    # Plot an mRNA track (mRNA + five_prime_UTR + three_prime_UTR) from gff
    # for only one mRNA transcript, names "gene"
    gffsmall <- gff %>%
        filter(Gene==gene)
    mRNA_plot <- 
      ggplot(data=gffsmall,
           aes(xmin=start,xmax=end)) +
        # geom_rect(data=filter(gffsmall,type=="mRNA"),
        #           aes(ymin=-1,ymax=1),fill="darkblue") +
        geom_rect(data=filter(gffsmall,type=="five_prime_UTR"),
                  aes(ymin=-4,ymax=4),fill="darkblue") +
        geom_rect(data=filter(gffsmall,type=="three_prime_UTR"),
                  aes(ymin=-4,ymax=4),fill="darkblue") +
        geom_rect(data=filter(gffsmall,type=="CDS"),
                  aes(ymin=-10,ymax=10),fill="darkblue") +
        theme_nothing()
        # theme(axis.text.x = element_text(colour="black"),
        #       axis.ticks.x=element_line(colour="black",size=0.5))
    if(!is.null(gff_extra)) {
        # select location of the mRNA features in gffsmall
        small_chrom <- gffsmall$chrom[1]
        small_strand <- gffsmall$strand[1]
        small_start <- min(gffsmall$start) - pad
        small_end   <- max(gffsmall$end) + pad
        # filter gff_extra for only completely overlapping features
        gff_extrasmall <- filter(gff_extra,
                                 chrom == small_chrom,
                                 strand == small_strand,
                                 start >= small_start,
                                 end <= small_end)
        mRNA_plot <- 
          mRNA_plot + 
          geom_rect(data=gff_extrasmall,
                    aes(ymin=-8,ymax=8),fill=colour_extra)
    }
    return(mRNA_plot)
}

# plot_mRNA(gff_H99,"CNAG_06125")

plot_mRNAbg <- function(bgbig,gff,gene,strandrev=FALSE,pad=100,ptitle=NULL,gff_extra=NULL,colour_extra="red") {
    gffmRNA <- gff %>%
        filter(Gene==gene)
    startr <- min(gffmRNA$start) - pad
    endr   <- max(gffmRNA$end) + pad
    bgplot <- plot_bg(bgbig,
                      chromr = gffmRNA$chrom[1],
                      startr = startr,
                      endr = endr,
                      strandrev = strandrev)
    mplot <- plot_mRNA(gff,gene,pad=pad,
                       gff_extra=gff_extra,colour_extra=colour_extra) 
    if (strandrev) {
        mplot <- mplot + 
            scale_x_reverse(expand=c(0,0),limits=c(endr,startr))
    } else {
        mplot <- mplot + 
            scale_x_continuous(expand=c(0,0),limits=c(startr,endr)) 
    }
    if (is.null(ptitle)) {
      ptitle <- gene
    }
    tplot <- ggdraw() + 
        draw_label(ptitle,fontface = 'bold',size = 11)
    plot_grid(tplot,mplot,bgplot,
              rel_heights = c(0.09,0.05,0.86),
              ncol=1,align="v",axis="lr")
}

plot_CDSbg <- function(bgbig,gff,gene,strandrev=FALSE,pad=1000,ptitle=NULL,gff_extra=NULL,colour_extra="red") {
    gffCDS <- gff %>%
        filter(Gene==gene,type=="CDS")
    startr <- gffCDS$start[1] - pad
    endr   <- gffCDS$end[1] + pad
    bgplot <- plot_bg(bgbig,
                      chromr = gffCDS$chrom[1],
                      startr = startr,
                      endr = endr,
                      strandrev = strandrev)
    mplot <- plot_mRNA(gff,gene,pad=pad,
                       gff_extra=gff_extra,colour_extra=colour_extra) 
    if (strandrev) {
        mplot <- mplot + 
            scale_x_reverse(expand=c(0,0),limits=c(endr,startr))
    } else {
        mplot <- mplot + 
            scale_x_continuous(expand=c(0,0),limits=c(startr,endr)) 
    }
    if (is.null(ptitle)) {
      ptitle <- gene
    }
    tplot <- ggdraw() + 
        draw_label(ptitle,fontface = 'bold',size=20)
    plot_grid(tplot,mplot,bgplot,
              rel_heights = c(0.05,0.03,0.92),
              ncol=1,align="v",axis="lr")
}

```


### Load bedgraphs

```{r load_bedgraph}
bedgraph_dir <- "../Ssd1_CRAC_demult_dedup_20190114_all/bedgraph_genomecov/"
bgall <- bind_rows(
    read_bedgraph(paste0(bedgraph_dir,"20190114_Ssd1_CRAC_trimmed_NNNGTGAGC_SSD1_3_30_plus.bedgraph")) %>%
        mutate(Type="",Sample="A",strand="plus",strandm=1),
    read_bedgraph(paste0(bedgraph_dir,"20190114_Ssd1_CRAC_trimmed_NNNGTGAGC_SSD1_3_30_minus.bedgraph")) %>%
        mutate(Type="",Sample="A",strand="minus",strandm=-1),
    read_bedgraph(paste0(bedgraph_dir,"20190114_Ssd1_CRAC_trimmed_NNNTGGAGC_SSD1_4_30_plus.bedgraph")) %>%
        mutate(Type="",Sample="B",strand="plus",strandm=1),
    read_bedgraph(paste0(bedgraph_dir,"20190114_Ssd1_CRAC_trimmed_NNNTGGAGC_SSD1_4_30_minus.bedgraph")) %>%
        mutate(Type="",Sample="B",strand="minus",strandm=-1)
)
```

### Load genome feature file

```{r load_gff}
gff_Sc <- read_gff("../input_annotation/gff_ncRNAs_abundantverifiedmRNAparts.gff") %>%
  mutate(ID = str_extract(attr,"ID=[A-Za-z0-9-_]+") %>%
           str_remove("ID="), 
         Parent = str_extract(attr,"Parent=[A-Za-z0-9-_]+")%>%
           str_remove("Parent="),
         Gene = str_extract(Parent,"Y[A-Z0-9-]+"))

# gff_SEEmotifs <- read_gff("../CRAC_Analysis/Data/SEEmotifs_CNYTCNYT_upstreamflank_MEME_13Feb2020/SEEmotif_sacCer3_both.gff")
```

## Plot 8 top Ssd1 targets - figure 4E in the manuscript

SUN4, SIM1, UTH1, SCW4, CTS1, DSE2, CCW12, SRL1.

All oriented with 5'UTR on the left, plotted in chromosomal co-ordinates.

```{r plot_8genes,dependson=c("load_bedgraph","load_gff","plot_functions"),fig.height=6,fig.width=6,units="in"}


SUN4_bg <- 
  plot_mRNAbg(bgbig=filter(bgall,Sample %in% c("A","B"),strand=="plus"),
           gff=gff_Sc,gene="YNL066W", ptitle="SUN4",pad=200)

SIM1_bg <- 
  plot_mRNAbg(bgbig=filter(bgall,Sample %in% c("A","B"),strand=="plus"),
           gff=gff_Sc,gene="YIL123W", ptitle="SIM1",pad=200)
CTS1_bg <- 
  plot_mRNAbg(bgbig=filter(bgall,Sample %in% c("A","B"),strand=="minus"),
           gff=gff_Sc,gene="YLR286C", ptitle="CTS1",pad=200, strandrev=TRUE)

DSE2_bg <- 
  plot_mRNAbg(bgbig=filter(bgall,Sample %in% c("A","B"),strand=="plus"),
           gff=gff_Sc,gene="YHR143W", ptitle="DSE2",pad=200)

UTH1_bg <- 
  plot_mRNAbg(bgbig=filter(bgall,Sample %in% c("A","B"),strand=="plus"),
           gff=gff_Sc,gene="YKR042W", ptitle="UTH1",pad=200)
SCW4_bg <- 
  plot_mRNAbg(bgbig=filter(bgall,Sample %in% c("A","B"),strand=="minus"),
           gff=gff_Sc,gene="YGR279C", ptitle="SCW4",pad=200,strandrev=TRUE)

CCW12_bg <- 
  plot_mRNAbg(bgbig=filter(bgall,Sample %in% c("A","B"),strand=="minus"),
           gff=gff_Sc,gene="YLR110C", ptitle="CCW12",pad=200, strandrev=TRUE)

SRL1_bg <- 
  plot_mRNAbg(bgbig=filter(bgall,Sample %in% c("A","B"),strand=="plus"),
           gff=gff_Sc,gene="YOR247W", ptitle="SRL1",pad=300)

plot_8genes <- 
  plot_grid(SUN4_bg,SIM1_bg,UTH1_bg,SCW4_bg,CTS1_bg,DSE2_bg,CCW12_bg,SRL1_bg,
            ncol = 2)

plot_8genes

ggsave("../figure_out/bedgraph-plot_8genes.pdf",plot_8genes,
       height=6,width=6)
```


## Session Info

Package versions, etc., that ran this file.

```{r session_info,results="show",message=TRUE}
sessionInfo()
```
