---
title: "Compare Ssd1 target transcripts against Hogan 2008, Jansen 2009, Hose 2020"
author: "Edward Wallace, Edward.Wallace@ed.ac.uk"
date: "19 February 2021"
output: 
    html_document:
        toc: true
---

This analysis takes the list of transcripts enriched in Ssd1 binding in the CRAC dataset, and compares them with Ssd1-targeted transcripts from previous publications that do RIP-ChIP or RIP-seq:

* Hogan 2008 
* Jansen 2009
* Hose 2020

These publicly available data need to be downloaded to the `input_targets` directory, see `input_targets/README.md` for details.

This script also relies on lists of Ssd1-enriched transcripts from DESeq2 analysis produced by `deseq2_Ssd1_CRAC_vs_RNAseq.Rmd`, found in `results/Ssd1_enrichment_DeSeq2_30C_2x_padj0p05_genesonly.txt`
and `results/Ssd1_enrichment_DeSeq2_30C_2x_padj0p05_genesonly.txt`.

The script outputs
* `results/Ssd1_targets_common_Hogan_Jansen_Bayne30C.txt`

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.path="figure/compare_target_transcripts-")

library(tidyverse)
library(Biostrings)
library(here)
library(VennDiagram)
venn <- function(...) {
    venn.diagram(x=list(...),filename=NULL,margin=0.1) %>% grid.draw
}
```

# Lists of mRNAs enriched in Ssd1 binding

## Hogan 2008

Reading data from Hogan 2008 `https://doi.org/10.1371/journal.pbio.0060255.sd003`.

```{r Hogan_targets}
# hoganFDRsall <- 
#     here::here("data","Ssd1_targets","Hogan2008_FDR_Dataset_S2.txt") %>%
#     read_tsv(comment="#")
# 
# hoganFDRsall

ssd1_data_hogan2008 <-
  readxl::read_excel(path = "../input_targets/journal.pbio.0060255.sd003_Hogan2008_Dataset_S3.xls",
                     sheet = "Ssd1_sam",skip=1,
                     col_names = c("Gene",
                                   "Score",
                                   "Numerator",
                                   "Denominator",
                                   "FoldChange",
                                   "qvalue",
                                   "localfdr")
  )

ssd1_data_hogan2008

ssd1_targets_hogan2008 <- 
    ssd1_data_hogan2008 %>%
    dplyr::filter(stringr::str_detect(Gene,"_ORF"), 
                  localfdr < 1) %>%
    dplyr::pull(Gene) %>%
    stringr::str_remove("_.*") %>%
    unique
```

## Jansen 2009

Reading data from Jansen 2009.

Surprisingly, the gene systematic names in this table are not completely systematic.

```{r Jansen_targets}
ssd1_data_jansen2009 <-
    readxl::read_excel(path = "../input_targets/NIHMS158901-supplement-02.xls",
                       sheet=1)

ssd1_data_jansen2009 

# Here we have to take some effort to get systematic names for all targets
ssd1_genic_targets_jansen2009 <- 
    ssd1_data_jansen2009 %>%
    dplyr::filter(`inter/genic?`=="genic") %>%
    dplyr::pull(sys_id)

ssd1_intergenic_targets_jansen2009 <- 
    c(ssd1_data_jansen2009 %>%
          dplyr::filter(`inter/genic?`!="genic",`sys_id`==`left gene name`) %>%
          dplyr::pull(`left_gene sys id`),
      ssd1_data_jansen2009 %>%
          dplyr::filter(`inter/genic?`!="genic",`sys_id`==`right gene name`) %>%
          dplyr::pull(`right_gene sys id`)
    )

ssd1_targets_jansen2009 <- unique( c(ssd1_genic_targets_jansen2009,
                                     ssd1_intergenic_targets_jansen2009) )
```

## Hose 2020

```{r Hose_targets}
ssd1_data_hose2020 <-
    readxl::read_excel(path = "../input_targets/elife-52063-supp2-v2_Hose2020.xlsx",
                       range = "A2:E6715",
                       col_names = c("Gene",
                                     "Annotation",
                                     "Bound",
                                     "RNA_DE",
                                     "Prot_DE")
    )

ssd1_data_hose2020

ssd1_targets_hose2020 <- 
    ssd1_data_hose2020 %>%
    dplyr::filter(Bound=="yes") %>%
    dplyr::pull(Gene)
```

# Venn diagram of intersections

Ugly yet effective.

```{r venn_hogan_jansen_hose,dependson=c("Hogan_targets","Jansen_targets","Hose_targets"),fig.width=4,fig.height=3.8}
venn(Hogan=ssd1_targets_hogan2008,
     Jansen=ssd1_targets_jansen2009,
     Hose=ssd1_targets_hose2020)
```


## Bayne targets

From DESeq2 analysis.

```{r Bayne_targets}
ssd1_targets_bayne2020_30C <- 
  readr::read_lines("../results/Ssd1_enrichment_DeSeq2_30C_2x_padj0p05_genesonly.txt")

ssd1_targets_bayne2020_42C <- 
  readr::read_lines("../results/Ssd1_enrichment_DeSeq2_42C_2x_padj0p05_genesonly.txt")
```


# Venn diagram of intersections with all datasets

Ugly yet effective.

```{r venn_hogan_jansen_hose_bayne,dependson=c("Hogan_targets","Jansen_targets","Hose_targets","Bayne_targets"),fig.width=5,fig.height=3.8}
venn(Hogan2008=ssd1_targets_hogan2008,
     Jansen2009=ssd1_targets_jansen2009,
     Hose2020=ssd1_targets_hose2020,
     Bayne30C=ssd1_targets_bayne2020_30C,
     Bayne42C=ssd1_targets_bayne2020_42C)
```


## Smaller Venn diagrams of Hogan 2008 vs Jansen 2009 vs Bayne 30C

This is figure panel 4D of the manuscript.

```{r venn_hogan_jansen_bayne,dependson=c("Hogan_targets","Jansen_targets","Bayne_targets"),dev="svg",fig.width=2,fig.height=2}
venn(Hogan2008=ssd1_targets_hogan2008,
     Jansen2009=ssd1_targets_jansen2009,
     Bayne30C=ssd1_targets_bayne2020_30C)
```


## What are the genes common to these three datasets?

First make a gene-name data frame to show nice gene names:

```{r load_genenames,warning=FALSE,message=FALSE}
gene_name_df <- 
  read_tsv("../input_annotation/scer-mrna-protein-absolute-estimate.txt",comment="#") %>%
  select(Gene=orf,Name=gene)
```

Then find the intersection. We are taking 2x enriched targets with FDR 5%, in Bayne 2020 at either 30 or 42C.

```{r Ssd1_commontargets,dependson=c("Hogan_targets","Jansen_targets","Bayne_targets","load_genenames"),results="show"}
ssd1_targets_common <- 
    ssd1_targets_bayne2020_30C %>%
    intersect(ssd1_targets_hogan2008) %>%
    intersect(ssd1_targets_jansen2009)
    
dplyr::filter(gene_name_df,Gene %in% ssd1_targets_common)
```

## Write output file of common targets

```{r Ssd1_commontargets_write,dependson="Ssd1_commontargets"}
ssd1_targets_common_file <- "../results/Ssd1_targets_common_Hogan_Jansen_Bayne30C.txt"

readr::write_lines(c("# Ssd1_targets_common_Hogan_Jansen_Bayne30C.txt",
                     "# Common targets of Ssd1 across 30C datasets, output from compare_target_transcripts.Rmd",
                     "# This table is the result of intersecting hit tables from Hogan 2008, Jansen 2009, and Bayne 2021 (this paper) list in `Ssd1_enrichment_DeSeq2_30C_2x_padj0p05_genesonly.txt`",
                     "# Edward.Wallace@ed.ac.uk, 6 Feb 2021",
                     "# "),
                   ssd1_targets_common_file)

dplyr::filter(gene_name_df,Gene %in% ssd1_targets_common) %>%
  readr::write_tsv(ssd1_targets_common_file, col_names = TRUE, append = TRUE)
```

## Session Info

Package versions, etc., that ran this file.

```{r session_info,results="show",message=TRUE}
sessionInfo()
```